---
title: "Predictive Analysis IV"
author: "Riki Herliansyah"
date: "13/10/2021"
output: github_document
---

## One-shot experiment
This is one-shot experiment: it is the outcome produced by a single simulated data set.

```{r}
n <- 100
xbar <- mean(rnorm(n, mean=20, sd=5))
xbar
```

## Monte Carlo Simulation
```{r}
set.seed(17)
R <- 1000
xbar <- numeric()
for (i in 1:R) {
  xbar[i] <- mean(rnorm(n, mean=20, sd=5))
}
mean(xbar)
hist(xbar, breaks=20)
abline(v=mean(xbar), col="red")

```


The second simulation study where we have:

$$Y_t = 0.5Y_{t-1} - 0.5Z_{t-1} + Z_t - 0.6X_t $$ 

where $\delta=0.5$, $\theta=0.5$, $\beta=-0.6$ and $Z_t \sim \text{N}(0,1)$.

```{r}
set.seed(17)
R <- 1000
deltahat <- numeric()
betahat <- numeric()
thetahat <- numeric()
for (i in 1:R) {
    Ut <- arima.sim(list(order=c(1,0,1), ar=0.5, ma=0.5), n=250)
    Xt <- rnorm(250, mean=10, sd=5)
    Yt <- Ut - 0.6*Xt
    fit <- arima(Yt, order = c(1,0,1), xreg = Xt, include.mean = FALSE)
    deltahat[i] <- fit$coef[1]
    thetahat[i] <- fit$coef[2]
    betahat[i] <- fit$coef[3]
}
mean(deltahat)
mean(thetahat)
mean(betahat)
par(mfrow=c(1,3))
hist(deltahat, breaks=20)
abline(v=mean(deltahat), col="red")
hist(thetahat, breaks=20)
abline(v=mean(thetahat), col="red")
hist(betahat, breaks=20)
abline(v=mean(betahat), col="red")

```

The following simulation study is to show how make predictions when there exists explanatory variables. The model used is the same as the previous model e.g., $$Y_t = \text{ARMA(1,1)} + \beta X_t$$. We split the data into training data set and testing data set for assessing the forecasts. 

```{r}
set.seed(17)
Ut <- arima.sim(list(order=c(1,0,1), ar=0.5, ma=0.5), n=250)
Xt <- rnorm(250, mean=10, sd=5)
Yt <- Ut - 0.6*Xt
Ytrain <- Yt[1:200]
Xtrain <- Xt[1:200]
    
fit <- arima(Ytrain, order = c(1,0,1), xreg = Xtrain, method = "ML", include.mean = FALSE)
fit
# the following to check if parameters are significant (TRUE is significant)
se.fit <- sqrt(diag(fit$var.coef)) 
abs(fit$coef[1]/se.fit[1]) > qnorm(0.975)
abs(fit$coef[2]/se.fit[2]) > qnorm(0.975)
abs(fit$coef[3]/se.fit[3]) > qnorm(0.975)

```
The following is used to forecast 50 points ahead. We need to provide extrapolation for $X_t$ for 50 points ahead as well, however. In practice, we might need to extrapolate this by assuming $10\%$ increase for example (scenario and sensitivity analysis).

```{r}
fit.fore <- predict(fit, newxreg=Xt[201:250], n.ahead=50)
ts.plot(as.ts(Yt), fit.fore$pred, col=c(1,2), lwd=c(1,2))
lines(fit.fore$pred + 1.96*fit.fore$se, col="blue", lty="dashed", lwd=1)
lines(fit.fore$pred - 1.96*fit.fore$se, col="blue", lty="dashed", lwd=1)
title("Forecasts")
legend("topleft",c( "Actual","Forecasts") , lty=c(1,1), lwd=c(2,2), col=c(1,2))
```


## Intervention Modeling for US Passenger Airlines
Note: you should put the data along with the code in the same folder.

```{r}
library(TSA)
data<-read.csv("USAirTraffic2017.csv",header=F)
USairmiles<-ts(data[,2],start=c(1996,1),frequency=12)
ts.plot(USairmiles,xlim=c(1996,2018))
abline(v=2001+8/12,lwd=2,col="red")   # 9/11 attacks indicators to end of nearest month (August)
abline(v=2007+9/12,lwd=2,col="green")   # GFC = ASX peaked in September 2007?
legend("bottomright",legend=c("Terrorist Attacks","GFC"),lty=c(1,1),col=c("red","green"))

```

The following is ACF and PACF of the data.
```{r}
par(mfrow=c(1,2))
acf(USairmiles)
pacf(USairmiles)
```

It seems that we need seasonally difference the data and then again differencing at lag 1 non-seasonal component.

```{r}
# Seasonally differenced data:
usair.d12<-diff(USairmiles,lag=12)
par(mfrow=c(1,3))
ts.plot(usair.d12)
abline(h=0)
acf(usair.d12)
pacf(usair.d12)

# Differenced and Seasonally Differenced Data
usair.d1d12 <- diff(diff(USairmiles, lag=12))
par(mfrow=c(1,3))
ts.plot(usair.d1d12)
abline(h=0)
acf(usair.d1d12)
pacf(usair.d1d12)

```

First, we will develop model using the airline model, ARIMA(0,1,1)$\times$(0,1,1)$_{12}$. 

```{r}
# 36 month ahead Forecasting using seasonal ARIMA model based on data to August 2001.
USairmiles.pre<-window(USairmiles,end=c(2005,8))
air.m0<-arima(log(USairmiles.pre),order=c(0,1,1),
              seasonal=list(order=c(0,1,1),period=12),method='ML')
air.m0

```

The next step is to check the residual plots.


```{r}

par(mfrow=c(1,1))
tsdiag(air.m0,gof.lag=40) # use the inbuilt 'tsdiag' function.
SWtest=shapiro.test(air.m0$residuals)
qqnorm(air.m0$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))

```

Finally, forecasting (even if we have a problem with non-normal residuals).

```{r}
air.m0.fore<-predict(air.m0,n.ahead=36)
ts.plot(log(USairmiles.pre),air.m0.fore$pred,col=c(1,4),lwd=c(1,2), ylim=c(17,18.2))
lines(air.m0.fore$pred+1.96*air.m0.fore$se,lty="dashed",col=2,lwd=1)
lines(air.m0.fore$pred-1.96*air.m0.fore$se,lty="dashed",col=2,lwd=1)
title("Log of US Airline PAX Miles Forecasts")
lines(window(log(USairmiles),start=c(2005,9),end=c(2008,8)),lwd=2)
legend("topleft",c("Forecasts", "Actual") , lty=c(1,1), lwd=c(2,2), col=c(4,1))
```

Now, I am going to show how to return the log forecasts into original values.

```{r}
#mean forecast
air.m0.meanfore<-exp(air.m0.fore$pred+0.5*air.m0.fore$se^2)
upper.log.m0 <- air.m0.fore$pred+1.96*air.m0.fore$se
lower.log.m0 <- air.m0.fore$pred-1.96*air.m0.fore$se

ts.plot(USairmiles.pre,air.m0.meanfore, col=c(1,4),lwd=c(1,2),ylim=c(2.5*10^7,7.5*10^7))
lines(exp(upper.log.m0+0.5*air.m0.fore$se^2),lty="dashed",col=2,lwd=1)
lines(exp(lower.log.m0+0.5*air.m0.fore$se^2),lty="dashed",col=2,lwd=1)
title("US Airline PAX Miles Forecasts")
lines(window(USairmiles,start=c(2005,9),end=c(2008,8)),lwd=2)
legend("topleft",c("Forecasts", "Actual") , lty=c(1,1), lwd=c(2,2), col=c(4,1))
#median forecast (remove hastag if you want to execute the code)
#air.m0.medfore<-exp(air.m0.fore$pred)
```

We are done with ARIMA model. Now, we are going to model the 9/11 attacks first as we show in the slide, that is we introduce an instantaneous major impact term based on a pulse and a decay term which eventually reduces to zero impact after an exponential decay.

$$ I_t = \omega_0 P_t (T) + \frac{\omega_1}{1-\delta B} P_t (T)$$

So, our model becomes:

$$ Y_t = I_t + U_t$$

where $U_t=\text{arima}(0,1,1)\times(0,1,1)_{12}$, the airline model, T= September 2001 and ðµ = backshift operator.

```{r}
# we create the pulse variable to model the short-term impact of 9/11 attacks
Nine11p=1*(seq(USairmiles.pre)==69) # pulse function at September 2001

air.all.m=arimax(log(USairmiles.pre),order=c(0,1,1),
                 seasonal=list(order=c(0,1,1), period=12),
                 xtransf=data.frame(Nine11p,Nine11p),
                 transfer=list(c(0,0),c(1,0)),method='ML')

air.all.m
```

As usual, we will check residual plot.

```{r}
par(mfrow=c(2,2))
plot(log(USairmiles.pre),ylab="log(airmiles)")
points(fitted(air.all.m))
abline(v=2001+9/12,lwd=2,col=3)    # 911 attacks
plot(air.all.m$residuals, type="h")
abline(v=2001+9/12,lwd=2,col=3)    # 911 attacks

acf(air.all.m$residuals)
SWtest=shapiro.test(air.all.m$residuals)
qqnorm(air.all.m$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))

```

The fitted intervention effect from the above model.

```{r}
omega_0=air.all.m$coef["Nine11p-MA0"]
omega_1=air.all.m$coef["Nine11p.1-MA0"]
delta=air.all.m$coef["Nine11p.1-AR1"]

AttackImpact<-ts(omega_0*Nine11p+omega_1*filter(Nine11p,filter=delta,method='recursive',side=1),
                 frequency =12, start=1996)
par(mfrow=c(1,1))
plot(AttackImpact,ylab='9/11 Impact', type='h'); abline(h=0)
```


Next, we will add a long-term effect:

$$ I_t = \omega_0 P_t (T) + \frac{\omega_1}{1-\delta B} P_t (T) + \omega_2 S_t(T)$$

After running the model, you will find that $\omega_0$ is not significant (try this out yourself). Since $\omega_0 is not significant, so our model becomes:

$$ Y_t = \frac{\omega_1}{1-\delta B} P_t (T) + \omega_2 S_t(T) + U_t + O_t$$

where $ð‘ˆ_ð‘¡$=arima(0,1,1)$\times$(0,1,1)$_{12}$, the airline model. 
An outlier (ð‘‚_ð‘¡) is also detected and modelled.

```{r}
Nine11s<-c(rep(0,68),rep(1,length(Nine11p)-68))  # Step Function at September 2001

air.m2=arimax(log(USairmiles.pre),order=c(0,1,1),
              seasonal=list(order=c(0,1,1), period=12),
              xtransf=data.frame(Nine11p,Nine11p, Nine11s),
              transfer=list(c(0,0),c(1,0),c(0,0)),method='ML')

air.m2

par(mfrow=c(2,2))
plot(log(USairmiles.pre),ylab="log(airmiles)")
points(fitted(air.m2))
plot(air.m2$residuals, type="h")
SWtest=shapiro.test(air.m2$residuals)
qqnorm(air.m2$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))


```

After checking the residual plot, we found an outlier and add an outlier dummy for the single positive outlier.

```{r}
#locating the positive outlier and defining a dummy variable for it
ndata=length(air.m2$residuals)
out=rep(0,ndata)
out[air.m2$residuals==max(air.m2$residuals)]=1
```

Fitting the model again with the inclusion an outlier dummy.

```{r}
air.m3=arimax(log(USairmiles.pre),order=c(0,1,1),
              seasonal=list(order=c(0,1,1),period=12),
              xtransf=data.frame(Nine11p, Nine11s,out),
              transfer=list(  c(1,0),   c(0,0),   c(0,0)),method='ML')
air.m3

par(mfrow=c(2,2))
plot(log(USairmiles.pre),ylab="log(airmiles)")
points(fitted(air.m3))
plot(air.m3$residuals, type="h")
acf(air.m3$residuals)
SWtest=shapiro.test(air.m3$residuals)
qqnorm(air.m3$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))

```

We will now draw the plot to see the impact of 9/11 atacks both a short-term and a long-term effect.

```{r}
omega_1=air.m3$coef["Nine11p-MA0"]
delta=air.m3$coef["Nine11p-AR1"]
omega_2=air.m3$coef["Nine11s-MA0"]

InterventionImpacts<-ts(omega_2*Nine11s
                        +omega_1*filter(Nine11p,filter=delta,method='recursive',side=1),
                        frequency =12, start=1996)
par(mfrow=c(1,1))
plot(InterventionImpacts,ylab='Impact', type='h'); abline(h=0)
```

Next, I want to show how we make future predictions using intervention modeling.

```{r}
# First, we need to define the regressor (explanatory variables)

Nine11p<-1*(seq(USairmiles.pre)==69) # pulse function at September 2011
Nine11Decay<-c(rep(0,68),delta^(0:(length(Nine11p)-69)))  # delta from Intervention Model air.m3
Nine11step<-c(rep(0,68),rep(1,length(Nine11p)-68))

#this is the effects that we are trying to include
par(mfrow=c(1,3))
ts.plot(Nine11step, col="blue", ylab="St(T)")
abline(v=69)
ts.plot(Nine11Decay, col="blue", ylab="Pt*(T)")
abline(v=69)
ts.plot(out, col="blue", ylab="Ot(T)")
abline(v=which.max(out))

#we combine them together
X=model.matrix(USairmiles.pre~ -1+   # -1 ensures no intercept is included
                 Nine11step+Nine11Decay+out)

N=length(log(USairmiles.pre))
L<-36 #(number of future predictions)
newNine11step<-rep(1,L)  # permanent change stays at 1.
newDecaynine11<-delta^(N+1:L-69)  # Effectively zero but put here to show general idea
outL <- rep(0,L)
#this is new values for our explanatory variables for 36 months ahead.
newxreg<-cbind(newNine11step,newDecaynine11,outL)
colnames(newxreg)<-colnames(X)
```

Let's make forecasts.

```{r}
model9.6.fixDecay<-arima(log(USairmiles.pre),order=c(0,1,1),xreg=X,
                         seasonal=list(order=c(1,1,1),period=12),method="ML")

model9.6.fixDecay   # almost identical (rounding) to results (point estimates and SE's)

#the following is the forecasts
model9.6.fixDecay.fore<-predict(model9.6.fixDecay,newxreg=newxreg,n.ahead=L)

#always check the residuals if we are happy with it
par(mfrow=c(2,2))
plot(log(USairmiles.pre),ylab="log(airmiles)")
points(fitted(model9.6.fixDecay))
plot(model9.6.fixDecay$residuals, type="h")
acf(model9.6.fixDecay$residuals)
SWtest=shapiro.test(model9.6.fixDecay$residuals)
qqnorm(model9.6.fixDecay$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))

```

Finally, visualise them.

```{r}
#mean forecast
model9.meanfore<-exp(model9.6.fixDecay.fore$pred+0.5*air.m0.fore$se^2)
upper.log <- model9.6.fixDecay.fore$pred+1.96*model9.6.fixDecay.fore$se
lower.log <- model9.6.fixDecay.fore$pred-1.96*model9.6.fixDecay.fore$se

par(mfrow=c(1,1))
ts.plot(log(USairmiles.pre), model9.6.fixDecay.fore$pred, air.m0.fore$pred, col=c(1,2,3),lwd=c(1,2), ylim=c(17,18.2))
lines(model9.6.fixDecay.fore$pred+1.96*model9.6.fixDecay.fore$se,lty="dashed",col=2,lwd=1)
lines(model9.6.fixDecay.fore$pred-1.96*model9.6.fixDecay.fore$se,lty="dashed",col=2,lwd=1)
lines(air.m0.fore$pred+1.96*air.m0.fore$se,lty="dashed",col=3,lwd=1)
lines(air.m0.fore$pred-1.96*air.m0.fore$se,lty="dashed",col=3,lwd=1)
title("US Airline PAX Miles Forecasts")
lines(window(log(USairmiles),start=c(2005,9),end=c(2008,8)),lwd=2)
legend("topleft",c( "Actual","Forecasts:IM","Forecasts:ARIMA") , lty=c(1,1), lwd=c(2,2), col=c(1,2,3))

```

This is the origin values.

```{r}
par(mfrow=c(1,1))
ts.plot(USairmiles.pre, model9.meanfore, air.m0.meanfore, col=c(1,2,3),lwd=c(1,2), ylim=c(2.5*10^7,7.5*10^7))
lines(exp(upper.log+0.5*model9.6.fixDecay.fore$se^2),lty="dashed",col=2,lwd=1)
lines(exp(lower.log+0.5*model9.6.fixDecay.fore$se^2),lty="dashed",col=2,lwd=1)
lines(exp(upper.log.m0+0.5*air.m0.fore$se^2),lty="dashed",col=3,lwd=1)
lines(exp(lower.log.m0+0.5*air.m0.fore$se^2),lty="dashed",col=3,lwd=1)
title("US Airline PAX Miles Forecasts")
lines(window(USairmiles,start=c(2005,9),end=c(2008,8)),lwd=2)
legend("topleft",c( "Actual","Forecasts:IM","Forecasts:ARIMA") , lty=c(1,1), lwd=c(2,2), col=c(1,2,3))

```

Now, we are going to combine both impacts into the model, 9/11 attacks and the 2008 GFC. First, we need to create the pulse (short-term) and the step (long-term) indicator variables for both impacts. 

```{r}
# 9/11 attacks
airmiles.all<-window(USairmiles,end=c(2014,12))
Nine11p=1*(seq(airmiles.all)==69) # pulse function at September 2001
Nine11s=c(rep(0,68),rep(1,length(Nine11p)-68))  # Step Function at September 2001#

# GFC
GFCp=1*(seq(airmiles.all)==153) # pulse function at September 2008
GFCs=c(rep(0,153),rep(1,length(GFCp)-153))  # Step Function at September 2008
```

Next is to develop the model.

```{r}
air.all=arimax(log(airmiles.all),order=c(0,1,1),
                               seasonal=list(order=c(0,1,2), period=12),
                               xtransf=data.frame(Nine11p, Nine11s,GFCp,GFCs),
                               transfer=list(c(1,0),c(0,0),c(1,0),c(0,0)),method='ML')

air.all
#remove outlier
ndata=length(air.all$residuals)
out.all=rep(0,ndata)
out.all[air.all$residuals==max(air.all$residuals)]=1

#remodel the data with ourlier now
air.all.m.maGFCresponse=arimax(log(airmiles.all),order=c(0,1,1),
                               seasonal=list(order=c(0,1,2), period=12),
                               xtransf=data.frame(Nine11p, Nine11s,GFCp,GFCs,out.all),
                               transfer=list(c(1,0),c(0,0),c(1,0),c(0,0),c(0,0)),method='ML')

air.all.m.maGFCresponse

```

Draw the effects as we did before.

```{r}
names(air.all.m.maGFCresponse$coef)

omega_1=air.all.m.maGFCresponse$coef["Nine11p-MA0"]
delta=air.all.m.maGFCresponse$coef["Nine11p-AR1"]
omega_2=air.all.m.maGFCresponse$coef["Nine11s-MA0"]

omega_1.dash=air.all.m.maGFCresponse$coef["GFCp-MA0"]
delta.dash=air.all.m.maGFCresponse$coef["GFCp-AR1"]
omega_2.dash=air.all.m.maGFCresponse$coef["GFCs-MA0"]

InterventionImpacts<-ts(omega_2*Nine11s
                        +omega_1*filter(Nine11p,filter=delta,method='recursive',side=1)
                        +omega_1.dash*filter(GFCp,filter=delta.dash,method='recursive',side=1)
                        +omega_2.dash*GFCs,
                        frequency =12, start=1996)
par(mfrow=c(1,1))
plot(InterventionImpacts,ylab='Impact', type='h'); abline(h=0)
```

Check the residuals and we are happy with the plots :).


```{r}
par(mfrow=c(2,2))
plot(log(airmiles.all),ylab="log(airmiles)")
points(fitted(air.all.m.maGFCresponse))
plot(air.all.m.maGFCresponse$residuals, type="h")
acf(air.all.m.maGFCresponse$residuals)
SWtest=shapiro.test(air.all.m.maGFCresponse$residuals)
qqnorm(air.all.m.maGFCresponse$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))
```

Finally, forecasts and visualisation following the previous steps.

```{r}
Nine11p<-1*(seq(airmiles.all)==69) # pulse function at September 2011
Nine11Decay<-c(rep(0,68),delta^(0:(length(Nine11p)-69)))  # delta from Intervention Model air.m3
Nine11step<-c(rep(0,68),rep(1,length(Nine11p)-68))

# GFC
GFCp=1*(seq(airmiles.all)==153) # pulse function at September 2008
GFCpdecay=c(rep(0,152),delta.dash^(0:(length(GFCp)-153))) # pulse function at September 2008
GFCs=c(rep(0,153),rep(1,length(GFCp)-153))  # Step Function at September 2008

X=model.matrix(airmiles.all~ -1+   # -1 ensures no intercept is included
                 Nine11step + Nine11Decay +
                 GFCs + GFCpdecay + out.all)

N=length(log(airmiles.all))
# Get actual values from from start t=1 to t= N+L 
L<-36
newNine11step<-rep(1,L)  # permanent change stays at 1.
newDecaynine11<-delta^(N+1:L-69)  # Effectively zero but put here to show general idea
newDecayGFC <- delta.dash^(N+1:L-153)
outL <- rep(0,L)

newxreg<-cbind(newNine11step,newDecaynine11,newNine11step,newDecayGFC,outL)
colnames(newxreg)<-colnames(X)

model9.6.fixDecay<-arima(log(airmiles.all),order=c(0,1,1),xreg=X,
                         seasonal=list(order=c(0,1,2),period=12),method="ML")

model9.6.fixDecay   # almost identical (rounding) to results (point estimates and SE's)

model9.6.fixDecay.fore<-predict(model9.6.fixDecay,newxreg=newxreg,n.ahead=L)
par(mfrow=c(2,2))
plot(log(airmiles.all),ylab="log(airmiles)")
points(fitted(model9.6.fixDecay))
plot(air.m3$residuals, type="h")
acf(model9.6.fixDecay$residuals)
SWtest=shapiro.test(model9.6.fixDecay$residuals)
qqnorm(model9.6.fixDecay$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))

#for comparison, we fit the airline model only
air.m0<-arima(log(airmiles.all),order=c(0,1,1),
              seasonal=list(order=c(0,1,2),period=12),method='ML')
air.m0
par(mfrow=c(1,1))
tsdiag(air.m0,gof.lag=40) # use the inbuilt 'tsdiag' function.
#title("Residual Diagnostics for ARIMA(0,1,1)x(0,1,1)12 model for US passenger miles",outer=TRUE)
SWtest=shapiro.test(air.m0$residuals)
qqnorm(air.m0$residuals, main=paste("Normal Q-Q plot \n SW test P-val=", as.character(round(SWtest$p.value,4))))

air.m0.fore<-predict(air.m0,n.ahead=36)
par(mfrow=c(1,1))
ts.plot(log(airmiles.all),model9.6.fixDecay.fore$pred, air.m0.fore$pred, col=c(1,2,3),lwd=c(1,2), ylim=c(17.1,18.2))
lines(model9.6.fixDecay.fore$pred+1.96*model9.6.fixDecay.fore$se,lty="dashed",col=2,lwd=1)
lines(model9.6.fixDecay.fore$pred-1.96*model9.6.fixDecay.fore$se,lty="dashed",col=2,lwd=1)
lines(air.m0.fore$pred+1.96*air.m0.fore$se,lty="dashed",col=3,lwd=1)
lines(air.m0.fore$pred-1.96*air.m0.fore$se,lty="dashed",col=3,lwd=1)
title("US Airline PAX Miles Forecasts")
lines(window(log(USairmiles),start=c(2015,1),end=c(2017,1)),lwd=1)
legend("topleft",c( "Actual","Forecasts:IM","Forecasts:ARIMA") , lty=c(1,1), lwd=c(2,2), col=c(1,2,3))


```



This is the origin values.

```{r}

#mean forecast
model9.meanfore<-exp(model9.6.fixDecay.fore$pred+0.5*air.m0.fore$se^2)
upper.log <- model9.6.fixDecay.fore$pred+1.96*model9.6.fixDecay.fore$se
lower.log <- model9.6.fixDecay.fore$pred-1.96*model9.6.fixDecay.fore$se

air.m0.meanfore<-exp(air.m0.fore$pred+0.5*air.m0.fore$se^2)
upper.log.m0 <- air.m0.fore$pred+1.96*air.m0.fore$se
lower.log.m0 <- air.m0.fore$pred-1.96*air.m0.fore$se

par(mfrow=c(1,1))
ts.plot(airmiles.all, model9.meanfore, air.m0.meanfore, col=c(1,2,3),lwd=c(1,2), ylim=c(2.5*10^7,7.5*10^7))
lines(exp(upper.log+0.5*model9.6.fixDecay.fore$se^2),lty="dashed",col=2,lwd=1)
lines(exp(lower.log+0.5*model9.6.fixDecay.fore$se^2),lty="dashed",col=2,lwd=1)
lines(exp(upper.log.m0+0.5*air.m0.fore$se^2),lty="dashed",col=3,lwd=1)
lines(exp(lower.log.m0+0.5*air.m0.fore$se^2),lty="dashed",col=3,lwd=1)
title("US Airline PAX Miles Forecasts")
lines(window(USairmiles,start=c(2015,1),end=c(2017,1)),lwd=1)
legend("topleft",c( "Actual","Forecasts:IM","Forecasts:ARIMA") , lty=c(1,1), lwd=c(2,2), col=c(1,2,3))
```